Search-Based Test Data Generation for JavaScript Functions that Interact with the DOM

### 2018-08-14 doc/issre2018/conclusions.tex
Prepared camera ready

In this paper we have introduced a novel JS unit testing framework, called \emph{JEDI}, which uses search-based techniques for generating test data. Notably, our testing framework is able to generate an arbitrary DOM input, which is syntactically valid and compliant with the recent HTML5 specification. We propose a novel test generation algorithm, called ``genetic with restart'', combining the benefits of systematic fitness improvement with the genetic search and the ability of escaping plateau by restarting search with a new target. Based on the case studies found in two related works~\emph{Confix}~\cite{amin:ase15} and \emph{TAJS}~\cite{dom2011}, we conduced an empirical validation of our framework and performed a significance study of the results. The validation has shown the effectiveness of the framework in covering target branches with a reasonable efficiency.

>>>>

In this paper we have introduced a novel search-based JS unit testing framework, called \emph{JEDI}. Notably, our framework is able to generate an arbitrary DOM inputs that are syntactically valid and compliant with the recent HTML5 specification. We present a test generation algorithm, called ``genetic with restart'', which is capable of escaping plateaus by concretizing the search target. Based on the case studies found in two related works~\emph{Confix}~\cite{amin:ase15} and \emph{TAJS}~\cite{dom2011}, we conduced an empirical validation followed by the significance study, which has confirmed the effectiveness and efficiency of our framework in branch coverage testing.

### 2018-08-14 doc/issre2018/evaluation.tex
Prepared camera ready

In order to evaluate the feasibility of our test framework, we conducted an empirical study on a set of subjects (see Section~\ref{sub.sec.case.studies}). All experiments were submitted as a batch mode job on a cluster machine with 32 HT Cores and 256 GB RAM running under Scientific Linux 7.3. We have created three different instances of the JS testing framework --- each one is based on a slight variation of the underlying genetic algorithm. The first version, which plays the role of our baseline, is \emph{Random} ($\Random$). Instead of applying crossover and mutation to evolve the population, this algorithm regenerates the new population randomly from scratch. It is important to note  that the random version also benefits from both preliminary static analysis and dynamic constant propagation as part of the fitness evaluation. The second version is \emph{Genetic}~($\Genetic$) --- it literally follows all of the steps prescribed by the genetic framework in Figure~\ref{fig.framework.architect}. The last version is \emph{Genetic with Restart} ($\RGenetic$). It is based on the genetic Algorithm~\ref{alg.gen} where the search is restarted after the archives are converged within a certain history window. On the one hand, the smaller window size forces the GA to converge faster and quickly discovers stagnation. On the other hand, the small window can lead to premature archive convergence, which prevents the right genetic combination from being triggered. We have experimented with the windows of size 5 and 10. But since the latter one ($\RGenetic$) showed the best performance over all, we chose to report only that case here. Table~\ref{tbl.gen.config} summarizes the configuration parameters for all three versions of the testing framework. The initial setup for genetic algorithm is similar to other frameworks, e.g. \emph{EvoSuite}~\cite{fraser2011evosuite}.

@@

To study the effectiveness and efficiency of our test generation framework, we exercised each algorithm ($\Random$, $\Genetic$ and $\RGenetic$) against all branches of the functions in Table~\ref{tbl.case.studies}. Due to the probabilistic nature of the test generation process, the experiments have been repeated 50 times to achieve statistical significance~\cite{arcuri2011practical}. Table~\ref{tbl.stats} summarizes the evaluation results. For each algorithm, we computed the following \emph{medians} per branch: the number of iterations~($i$), time in seconds~($t$), and their ratio~($t/i$). In the table, these values are grouped into three columns $m_{\Random}$, $m_{\Genetic}$ and $m_{\RGenetic}$ one for each algorithm. The next group of columns $t_{\Random}/t_{\Genetic}$, $t_{\Random}/t_{\RGenetic}$ and $t_{\Genetic}/t_{\RGenetic}$ presents the relative performance of each pair of the testing algorithm $(\Random,\Genetic)$, $(\Random,\RGenetic)$, and $(\Genetic,\RGenetic)$ respectively. To the right of the function name, we also present accumulated statistics for the function as a whole\footnote{because of the missing branches, the function's summary may not be equal to the sum of the branches, e.g. see the \emph{toggleInfo} function}

@@

\textbf{RQ1 (effectiveness)}: To answer this question, we measure the branch coverage achieved by each test generation algorithm. Given a fixed limit on the number of test iterations, 200 in our case, a search algorithm cannot guarantee finding a solution. The underlined values in Table~\ref{tbl.stats} like $\underline{200}$ indicate branches which were not covered by the respective algorithms. The main reason for the random generation~($\Random$) to fail is the large size of the search space and the lack of any feedback in the generation process, e.g. branch $(10,11)$ of the \emph{isGameFinished} function. In the case of the genetic approach~($\Genetic$), it  may get stuck at a plateau because the fitness function does not provide sufficient guidance, whereas $\Random$ could still escape the plateau by sheer luck. We encountered three such cases in our study: branches $(10,11)$, $(47,48)$ and $(45,45)$ of the function \emph{doPaddlePower}, \emph{initBricks} and \emph{do_draw}, respectively. They motivated us to introduce the genetic algorithm with restart ($\RGenetic$) which combines the strengths of both $\Random$ and $\Genetic$. It leverages the genetic search while the fitness continues to improve, and restarts the search with a new target, when the fitness suboptimally converges. As our evaluation has shown, $\RGenetic$ not only managed to outperform $\Random$ and cover the three problematic branches above, but it has generally improved the results of $\Genetic$, e.g. branch $(69,71)$ of the \emph{initBricks} function.\\
\framebox{
  \parbox[t][1.1cm]{0.95\linewidth}{
    Across all subjects, the $\RGenetic$ algorithm achieved 100\% branch coverage, with $\Genetic$ in the second place with 95\% coverage, and, finally, $\Random$ with  63\% coverage.
  }
}

@@

\textbf{RQ4 (comparison):} We evaluated \Confix\ as the only available alternative for the generation of DOM fixtures. In \cite{amin:ase15} the authors have proposed three strategies for test generation based on:
\begin{enumerate*}[label=(\roman*)]
  \item \emph{Confix} entirely,
  \item \emph{Jalangi}~\cite{sen2013jalangi} combination, and
  \item semi-manual approach.
\end{enumerate*}
Even though the third strategy outperformed pure \Confix by 19\%, it is not suitable for an objective comparison with a fully automated technique, like \Jedi. On the other hand, the combination with \Jalangi could be used for the evaluation, but it only improved branch coverage by 2\%, and the \Confix version available online~\cite{confixgit} has no actual integration with \Jalangi. Thus we had to fall back on plain \Confix test generation; results are present in Figure~\ref{tbl.confix.compare}. In comparison with \Jedi, \Confix does not search for a test input in terms of a \emph{strong} coverage, i.e. it only tries to reach a target branch neglecting the normal termination. But even in terms of \emph{weak} coverage, it showed the complete coverage in 2 out 10 cases.\\
\framebox{
  \parbox[t][0.65cm]{0.95\linewidth}{
Although, \Confix performs much faster (under 10 seconds), it is unable to cover hard branches, which \Jedi can.
  }
}

\textbf{Threats to Validity:}
The \emph{construct validity} threat comes from the choice of measurements: branch coverage and execution time. Although, the branch coverage is a common criterion used in practice, the actual fault finding capability is a stronger indication of testing adequacy. Reaching a target branch is only part of the problem: we also need an oracle to discriminate the program's output. In this paper, we rely on so called ``natural'' oracles leading to exceptions. Using a stronger oracle for validation is future work. Another practical aspect of the test generation is the test case size, since it can both complicate the readability and increase the execution time. To mitigate this issue we configured the GA to start evolution with the candidates of small sizes, which helps to prevent data explosion. Additional measure that could be applied is a minimization procedure, aka \emph{hierarchical delta debugging}~\cite{misherghi2006hdd} and its variations~\cite{herfert2017automatically}, to the GA result.

>>>>

In order to evaluate the feasibility of our test framework, we conducted an empirical study on a set of subjects (see Section~\ref{sub.sec.case.studies}). All experiments were submitted as a batch mode job on a cluster machine with 32 HT Cores and 256 GB RAM running under Scientific Linux 7.3. We have created three different instances of the JS testing framework --- each one is based on a slight variation of the underlying GA. The first version, which plays the role of our baseline, is \emph{Random} ($\Random$). Instead of applying crossover and mutation to evolve the population, this algorithm regenerates the new population randomly from scratch. It is important to note  that the random version also benefits from both preliminary static analysis and dynamic constant propagation as part of the fitness evaluation. The second version is \emph{Genetic}~($\Genetic$) --- it literally follows all of the steps prescribed by the genetic framework in Figure~\ref{fig.framework.architect}. The last version is \emph{Genetic with Restart} ($\RGenetic$). It is based on Algorithm~\ref{alg.gen} where the search is restarted after the archives are converged within a certain history window. On the one hand, the smaller window size forces the GA to converge faster and quickly discovers stagnation. On the other hand, the small window can lead to premature archive convergence, which prevents the right genetic combination from being triggered. We have experimented with the windows of size 5 and 10. But since the latter one ($\RGenetic$) showed the best performance over all, we chose to report only that case here. Table~\ref{tbl.gen.config} summarizes the configuration parameters for all three versions of the testing framework. The initial setup for the GA is similar to other frameworks, e.g. \emph{EvoSuite}~\cite{fraser2011evosuite}.

@@

To study the effectiveness and efficiency of our test generation framework, we exercised each algorithm ($\Random$, $\Genetic$ and $\RGenetic$) against all branches of the functions in Table~\ref{tbl.case.studies}. Due to the probabilistic nature of the test generation process~\cite{arcuri2011practical}, the experiments have been repeated 50 times to achieve statistical significance. Table~\ref{tbl.stats} summarizes the evaluation results. For each algorithm, we computed the following \emph{medians} per branch: the number of iterations~($i$), time in seconds~($t$), and their ratio~($t/i$). In the table, these values are grouped into three columns $m_{\Random}$, $m_{\Genetic}$ and $m_{\RGenetic}$ one for each algorithm. The next group of columns $t_{\Random}/t_{\Genetic}$, $t_{\Random}/t_{\RGenetic}$ and $t_{\Genetic}/t_{\RGenetic}$ presents the relative performance of each pair of the testing algorithm $(\Random,\Genetic)$, $(\Random,\RGenetic)$, and $(\Genetic,\RGenetic)$ respectively. To the right of the function name, we also present accumulated statistics\footnote{because of the missing branches, the function's summary may not be equal to the sum of the branches, e.g. see the \emph{toggleInfo} function}.

@@

\textbf{RQ1 (effectiveness)}: To answer this question, we measure the branch coverage achieved by each test generation algorithm. Given a fixed limit on the number of test iterations, 200 in our case, a search algorithm cannot guarantee finding a solution. The underlined values in Table~\ref{tbl.stats} like $\underline{200}$ indicate branches which were not covered by the respective algorithms. The main reason for the random generation~($\Random$) to fail is the large size of the search space and the lack of any feedback in the generation process, e.g. branch $(10,11)$ of the \emph{isGameFinished} function. In the case of the genetic approach~($\Genetic$), it  may get stuck at a plateau because the fitness function does not provide sufficient guidance, whereas $\Random$ could still escape the plateau by sheer luck. We encountered three such cases in our study: branches $(10,11)$, $(47,48)$ and $(45,45)$ of the function \emph{doPaddlePower}, \emph{initBricks} and \emph{do_draw}, respectively. They motivated us to introduce the GA with restart ($\RGenetic$) which combines the strengths of both $\Random$ and $\Genetic$. It leverages the genetic search while the fitness continues to improve, and restarts the search with a new target, when the fitness suboptimally converges. As our evaluation has shown, $\RGenetic$ not only managed to outperform $\Random$ and cover the three problematic branches above, but it has generally improved the results of $\Genetic$, e.g. branch $(69,71)$ of the \emph{initBricks} function.\\
\framebox{
  \parbox[t][1.1cm]{0.95\linewidth}{
    Across all subjects, the $\RGenetic$ algorithm achieved 100\% branch coverage, with $\Genetic$ in the second place with 95\% coverage, and, finally, $\Random$ with  63\% coverage.
  }
}

@@

\textbf{RQ4 (comparison):} We evaluate \Confix~\cite{amin:ase15} as the only available alternative for the generation of DOM fixtures. The authors suggest that \Confix\ achieves higher branch and statement coverage by being combined with \Jalangi~\cite{sen2013jalangi} --- a concolic test generation tool for JS --- which is able to produce test inputs for JS arguments of primitive types. However, the available online implementation of \Confix~\cite{confixgit} has no actual integration with \Jalangi. Thus we have to fall back on plain \Confix test generation.

Experimental results are presented in Figure~\ref{tbl.confix.compare}. We report both \emph{weak} (reach the target branch) and \emph{strong} (reach the branch and terminate normally) coverage. \Confix test generation is guided by the weak coverage in contrast with the strong coverage of \Jedi. In all experimental subjects except one (\jsID{isGameFinished}), \Confix poorly performs simply due to an inability to produce appropriate input arguments. However, in the case of \jsID{isGameFinished}, which requires no inputs, it still fails to reach deeply nested branches because of imprecise symbolic representation. In particular, it cannot interpret the expression on line~\ref{isGameFinished.if1.begin} in Figure~\ref{code.isGameFinished}:
%% following expression:
\begin{center}
\begin{lstlisting}[style=htmlcssjs,language=JavaScript,basicstyle={\footnotesize\ttfamily},backgroundcolor=\color{white},frame=none,numbers=none,xleftmargin={0.2cm}]
subDivs[no].className.indexOf('sudokuSquare') >= 0
\end{lstlisting}
\end{center}
This is not a principal constraint, but to be fully addressed, \Confix should model an extensive library of built-in JS functions. \Jedi mitigates this constraint by substituting observed string literals as conceivable class names, which are additionally evolved thanks to crossover and mutation operators. We admit, however, that some constructs, e.g. jQuery, are supported by Confix but not by Jedi.\\
\framebox{
  \parbox[t][1.5cm]{0.95\linewidth}{
In a nutshell, the choice between a concolic and search-based approach is a trade-off between a labour-intensive modelling effort and long execution time, respectively, where both can reinforce each other.
  }
}

\textbf{Threats to Validity:}
The \emph{construct validity} threat is due to the choice of measurements: branch coverage and execution time. Although, the branch coverage is a common criterion used in practice, the fault finding capability is a stronger indication of testing adequacy. Reaching a target branch is only part of the problem: we also need an oracle to discriminate the program's output. In this paper, we rely on so called ``natural'' oracles such as exceptions.
%Using a stronger oracle for validation is future work.
Another practical aspect of the test generation is the test case size, since it can both complicate the readability and increase the execution time. To mitigate this issue we configured the GA to start evolution with the candidates of small sizes, which helps to prevent data explosion. Additional measure that could be applied is a minimization procedure, aka \emph{hierarchical delta debugging}~\cite{misherghi2006hdd} and its variations~\cite{herfert2017automatically}, to the GA result.

### 2018-08-14 doc/issre2018/example.tex
Prepared camera ready

%% --------------------------------------------------------------------
\section{Motivating Example}
\label{sec.example}
%% --------------------------------------------------------------------

@@

Let us consider the JS function \texttt{isGameFinished} in Figure~\ref{code.isGameFinished} taken from the web game \emph{sudoku}~\cite{sudoku}. This function checks if a candidate sudoku solution is valid. First, it locates a DOM element \texttt{obj} representing the game field (line~\ref{isGameFinished.getSudoku}). Then, it collects all children \emph{DIV} elements of \texttt{obj} (line~\ref{isGameFinished.getDivs}). Each \emph{DIV} corresponds to an input square of the game (line~\ref{isGameFinished.if1.begin}) and consists of two \emph{SPAN} tags. The first \emph{SPAN} contains the value entered by the player, whereas the second one (hidden by default) stores the expected (correct) value for that square. In the for-loop in lines~\ref{isGameFinished.inFor.begin}-\ref{isGameFinished.outFor}, we iterate through the whole collection of the input squares. Once, we encounter a square with two unequal values (line~\ref{isGameFinished.if2.begin}), we terminate the loop and call the game unfinished (line~\ref{isGameFinished.unfinished}), otherwise finished.

Suppose now that we would like to unit test this function. We aim to maximize the branch coverage as a common testing criterion~\cite{zhu1997software}. That is, for each individual branch of the function under test (FUT) \texttt{isGameFinished}, we need to construct a test input that covers the branch and leads to the normal termination of the FUT. A JS function can take \emph{explicit} input arguments (in our case none), but it can also accept \emph{implicit} arguments such as a DOM state. In general, in order to produce a complete test for our function we have to construct both an appropriate DOM and input arguments. The tests should be decoupled from the application's HTML. Moreover, such HTML could still be under active development. Thus, the developers have to complement test with DOM fixtures.

@@

Figure~\ref{fig.isGameFinished.tests} exhibits four tests that together provide the full branch coverage of the FUT. Let us, for example, take a look at how we can reach the target branch $(10,11)$. Line~\ref{isGameFinished.getSudoku} expects the DOM to have an element with an id \textquotesingle\texttt{sudoku}\textquotesingle. In order to enter the for-loop in line~\ref{isGameFinished.inFor.begin}, this element should have at least one child \emph{DIV} tag. The first if-condition in line~\ref{isGameFinished.if1.begin} requires each \emph{DIV} element to be of the class \textquotesingle\texttt{square}\textquotesingle\, and do not have a background color. The second if-condition (line~\ref{isGameFinished.if2.begin}) implies that the \emph{DIV} should consist of two \emph{SPAN} elements whose \emph{innerHtml} values are not equal. The resulting test \texttt{T4} in Figure~\ref{fig.isGameFinished.tests} meets all the above conditions.

As we have just witnessed, even for such a relatively simple example, the test data generation can be far from trivial procedure because it requires the deep understanding of the program semantics. In practice, web developers often have to deal with third-party code, including a large number of JS libraries freely available on GitHub, which are commonly untested and poorly documented. At the same time, unit testing is the most fundamental level of testing and the easiest to implement. It saves the code from unexpected regressions and often documents the desired behavior of a program. Thus, in industry unit testing has become a standard practice in web development.

>>>>

%%% --------------------------------------------------------------------
\section{Motivating Example}
\label{sec.example}
%% --------------------------------------------------------------------

@@

Let us consider the JS function \jsID{isGameFinished} in Figure~\ref{code.isGameFinished} taken from the web game \emph{sudoku}~\cite{sudoku}. This function checks if a candidate sudoku solution is valid. First, it locates a DOM element \jsID{obj} representing the game field (line~\ref{isGameFinished.getSudoku}). Then, it collects all children \emph{DIV} elements of \jsID{obj} (line~\ref{isGameFinished.getDivs}). Each \emph{DIV} corresponds to an input square of the game (line~\ref{isGameFinished.if1.begin}) and consists of two \emph{SPAN} tags. The first \emph{SPAN} contains the value entered by the player, whereas the second one (hidden by default) stores the expected (correct) value for that square. In the for-loop in lines~\ref{isGameFinished.inFor.begin}-\ref{isGameFinished.outFor}, we iterate through the whole collection of the input squares. Once, we encounter a square with two unequal values (line~\ref{isGameFinished.if2.begin}), we terminate the loop and call the game unfinished (line~\ref{isGameFinished.unfinished}), otherwise finished.

Suppose now that we would like to unit test this function. We aim to maximize the branch coverage as a common testing criterion~\cite{zhu1997software}. That is, for each individual branch of the function under test (FUT) \jsID{isGameFinished}, we need to construct a test input that covers the branch and leads to the normal termination of the FUT. A JS function can take \emph{explicit} input arguments (in our case none), but it can also accept \emph{implicit} arguments such as a DOM state. In general, in order to produce a complete test for our function we have to construct both an appropriate DOM and input arguments. The tests should be decoupled from the application's HTML. Moreover, such HTML could still be under active development. Thus, the developers have to complement test with DOM fixtures.

@@

Figure~\ref{fig.isGameFinished.tests} exhibits four tests that together provide the full branch coverage of the FUT. Let us, for example, take a look at how we can reach the target branch $(10,11)$. Line~\ref{isGameFinished.getSudoku} expects the DOM to have an element with the id \textquotesingle\jsID{sudoku}\textquotesingle. In order to enter the for-loop in line~\ref{isGameFinished.inFor.begin}, this element should have at least one child \emph{DIV} tag. The first if-condition in line~\ref{isGameFinished.if1.begin} requires each \emph{DIV} element to be of the class \textquotesingle\jsID{square}\textquotesingle\, and do not have a background color. The second if-condition (line~\ref{isGameFinished.if2.begin}) implies that the \emph{DIV} element should consist of two \emph{SPAN} elements whose \emph{innerHtml} values are not equal. The resulting test \jsID{T4} in Figure~\ref{fig.isGameFinished.tests} meets all the above conditions.

As we have just witnessed, even for such a relatively simple example, the test data generation can be far from trivial procedure because it requires the deep understanding of the program semantics. In practice, web developers often have to deal with third-party code, including a large number of JS libraries freely available on GitHub, which are commonly untested and poorly documented. At the same time, unit testing is the most fundamental and easiest to implement type of testing. It saves the code from unexpected regressions and documents the desired behavior of a program. Thus, in industry unit testing has become a standard development practice.
